---
title: "Automation of Tax Statistics"
author: "Planning and Research Department"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
    theme: united
    highlight: espresso
    toc: true
    toc_float: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Dear users, you are welcome to RRA data portal for tax statistics analysis and visualization. There are 5 ways to export or save the summarized data and this includes `Copy` to clipboard, `CSV`, `Excel`, `PDF` or `Print` directly.

```{r Instruction note, include=F}
# We have quite number of chunks in this script.
# Each chunck has its name on the header
# We used many pipe or chain functions `%>%` in this script
```


```{r library,include=FALSE}

# This chunk listed the R packeges that will be needed for the automation

pg=c('tidyverse','openxlsx','readxl','gridExtra','DT','plotly')
for (i in pg){
  library(i,character.only = T)
}
```

  
```{r Dataset from registration, include=F, cache=T,include=F}

# This is the registration clean up file from Harshil

hasil.file <- read.xlsx('Dataset/file from Hashil.xlsx')
```

#  Tax revenue by sector

```{r tax_rev_by_sector_2017_18,cache=T, dependson = "Hashil_data",include=F}

# Tax revenue payment collection data from BI

tax_rev1718=read.xlsx('Dataset/Tax revenue FY1718 from BI.xlsx',sheet = 1)

# We selected the columns that we needed for the analysis

tax_rev1718_S <- tax_rev1718 %>%select(1,2,18,25,28) %>% mutate_at(vars(5),~parse_number(.)) # ~parse_number is to convert column 28 to numeric data type

# We joined registration data to collection data to get ISIC

tax_rev_by_sector <-left_join(tax_rev1718_S,hasil.file,by='TIN')

# We selected the columns that we wanted

tax_rev_by_sector <- tax_rev_by_sector %>% select(1,2,3,4,5,7,8) 

# We addeded conditional statement to the isicdesc such that if isicdesc is empty and totalcustomstax>0, then we called isicdesc 'Occasional Importer' else we called it 'Other service'

tax_rev_by_sector$isicdesc= if_else(is.na(tax_rev_by_sector$isicdesc)& tax_rev_by_sector$Total.Customs.Tax>0,'Occasional Importers','Other service')

# Recoding ISIC by Sectors and aggregating by Total

tax_rev_by_sector=tax_rev_by_sector %>% mutate(sector=recode_factor(isic,
             'A'='Agriculture',
             'B'='Industry',
             'C'='Industry',
             'D'='Industry',
             'E'='Industry',
             'F'='Industry',
             .default = 'Services',
             .missing = 'Services')) %>% 
  group_by(sector) %>% summarise(`2017/18`=sum(Total,na.rm = T))
tax_revenue_by_sector_FY2017_18=tax_rev_by_sector
```

```{r tax_revenue_by_sector_FY2009_17,include=F}

# We loaded the analysis done before by Harshil for 2009 to 2017

tax_rev_FY2009_17=read.xlsx('Dataset/Tax revenue FY1718 from BI.xlsx',sheet = 2)

# Recoding ISIC by Sectors and aggregating by Total

tax_rev_by_sector_FY2009_17=tax_rev_FY2009_17 %>% mutate(sector=recode_factor(isic,
             'A'='Agriculture',
             'B'='Industry',
             'C'='Industry',
             'D'='Industry',
             'E'='Industry',
             'F'='Industry',
             .default = 'Services',
             .missing = 'Services')) %>% 
  group_by(sector) %>% summarise_if(is.numeric, sum, na.rm=T)
```


```{r tax_rev_by_sector_YB,include=F}

# Combining the analysis of FY 2018 to the FY2009_17

tax_rev_by_sector_YB <- bind_cols(tax_rev_by_sector_FY2009_17,tax_revenue_by_sector_FY2017_18[,2]) 

# Converting the revenue for all the years in RWF billion
tax_rev_by_sector_YB <- tax_rev_by_sector_YB %>% mutate_at(vars(2:10),~(round(./10^9,2)))

tax_rev_by_sector_YB <- tax_rev_by_sector_YB %>% rename(Sector=sector) 

tax_rev_by_sector_YB$Sector <-  as.character(tax_rev_by_sector_YB$Sector)

# Grand total calculation 

a=summarise_all(tax_rev_by_sector_YB[-1],sum);b=tibble(Sector='Total')

Total=bind_cols(b,a)

tax_rev_by_sector_YB=bind_rows(tax_rev_by_sector_YB,Total)
```

```{r tax_rev_by_sector_final,include=F}

# We manipulated the data by using Percentage from the BI data `tax_rev_by_sector_final` to the real collection data

# This is done in Excel sheet

actual_revenue_by.sector=tribble(
  ~`Sector`, ~`2009/10`, ~`2010/11`, ~`2011/12`, ~`2012/13`, ~`2013/14`, ~`2014/15`, ~`2015/16`, ~`2016/17`, ~`2017/18`,
            "Agriculture",      2.1,      3.6,      6.1,        7,      9.2,      9.2,     14.1,     11.6,     12.6,
           "Industry",     91.7,    116.1,    151.2,    163.3,    188.1,    191.8,    230.7,    252.4,    278.9,
           "Services",      283,    340.9,    398.7,    481.6,    566.2,    658.2,    741.8,    822.5,    942.9,
              "Total",    376.8,    460.6,      556,    651.9,    763.5,    859.2,    986.6,   1086.5,   1234.4)

# Percentage from BI data								

percentage=tribble(
  ~`Sector`, ~`2009/10`, ~`2010/11`, ~`2011/12`, ~`2012/13`, ~`2013/14`, ~`2014/15`, ~`2015/16`, ~`2016/17`, ~`2017/18`,
             "Agriculture",   "0.6%",   "0.8%",   "0.6%",   "0.9%",   "0.9%",   "0.9%",   "1.2%",   "1.1%",     "1%",
           "Industry",  "24.3%",  "25.2%",  "26.1%",  "24.6%",  "23.8%",  "23.9%",  "24.3%",  "23.2%",    "23%",
           "Services",  "75.1%",  "74.0%",  "73.3%",  "74.5%",  "75.3%",  "75.2%",  "74.6%",  "75.7%",    "76%")
```



```{r data table for actual Sector, echo=F}
# using `DT` Java library to get download page

datatable(actual_revenue_by.sector,caption = "Tax revenue by sector:", extensions = 'Buttons',
options = list(dom='Bfrtip',                                                buttons=c('copy', 'csv', 'excel', 'pdf', 'print')))
```



```{r data table for % Sector, echo=F}
# using `DT` Java library to get download page

datatable(percentage,caption = "Percentage tax revenue by sector:",extensions = 'Buttons',options = list(dom='Bfrtip', buttons=c('copy', 'csv', 'excel', 'pdf', 'print')))
```

# Visualization of Tax revenue by sector
```{r tax_rev_by_sector_YB_Chart,echo=F}
viz1=actual_revenue_by.sector[1:3,] %>%  gather(key = 'Year',value = 'Revenue',-`Sector`)

viz2=percentage[1:3,] %>% gather(key = 'Year',value = 'Percentage',-`Sector`)

chart.table=bind_cols(viz1,viz2[,3]) %>% mutate(Year=factor(Year))

g=ggplot(chart.table,aes(x=Year,y=Revenue, fill=Sector))+
  geom_bar(stat = 'identity')+theme_bw()+labs(x='Fiscal year',y='Tax revenue (Rwf bn)', fill='Sector', caption = "From RRA data", tag= 'Fig.1', title='Tax revenue by sector')+
  theme(legend.position = 'top',axis.title.y = element_text(size = 12, face='bold'), axis.title.x =element_text(size = 12, face='bold'))

ggplotly(g)
```

# Tax revenue by enterprise size

```{r tax revenue by tax enterprise size FY2017_18,include=F}
# Grouping tax payers into enterprise size

tax_rev1718_E= tax_rev1718%>%select(1,2,3,18,25,28) %>% mutate_at(vars(6),~parse_number(.))%>% rename(`Scale desc`=Scale.desc) %>% mutate_at(vars(3),~as_factor(.)) %>% mutate(`Scale desc`=fct_collapse(`Scale desc`,      Small=c("SMALL","NON-BUSINESS", "MICRO"),
     Medium="MEDIUM",
     Large="LARGE"))
tax_rev1718_E=tax_rev1718_E %>% mutate(`Scale desc`=fct_relevel(`Scale desc`,'Small',after = 0))

tax_rev1718_E <- tax_rev1718_E %>% group_by(`Scale desc`) %>% summarise(`2017/18`=sum(Total,na.rm = T))
```

```{r tax revenue by enterprise size FY2009_17, include=F}
# This data was from the analysis done before

tax_rev_2009_17_E <-  tribble(
  ~`Size`, ~`2009/10`, ~`2010/11`, ~`2011/12`, ~`2012/13`, ~`2013/14`, ~`2014/15`, ~`2015/16`, ~`2016/17`,
                "Small",     52011365645,     76194136968,    135613977049,    139069324669,    171907720828,    211232367056,    261142501500,    265021785376,
               "Medium",     60227375104,     77876912128,    104499830784,     92267200512,     93977206784,    101571870720,    107970920448,    139249172480,
                "Large",    248441716736,    309714518016,    412835577856,    433031479296,    509692411904,    569827262464,    620696109056,    680577269760)

```

```{r tax revenue by tax enterprise size_YB,include=F}
# Converting the revenue for all the years in RWF billion

tax_rev_by_enterprise_YB <- bind_cols(tax_rev_2009_17_E,tax_rev1718_E[,2]) 
tax_rev_by_enterprise_YB <- tax_rev_by_enterprise_YB %>% mutate_at(vars(2:10),~(round(./10^9,2)))

# tax_rev_by_sector_YB$ <-  as.character(tax_rev_by_sector_YB$`Revenue (Rwf bn)`)

# Calculating Grand Total
a=summarise_all(tax_rev_by_enterprise_YB[-1],sum);b=tibble(Size='Total')

Total=bind_cols(b,a)
tax_rev_by_enterprise_YB=bind_rows(tax_rev_by_enterprise_YB,Total)
```

```{r tax_rev_by_enterprisesize_final,include=F}

# We manipulated the data by using Percentage from the BI data `tax_rev_by_enterprise_YB` to the real collection data

# This is done in Excel sheet

actual_revenue_by.enterprise <- tribble(
  ~`Size`, ~`2009/10`, ~`2010/11`, ~`2011/12`, ~`2012/13`, ~`2013/14`, ~`2014/15`, ~`2015/16`, ~`2016/17`, ~`2017/18`,
  "Small",     54.3,     75.7,      6.1,        7,      9.2,      9.2,     14.1,    265.4,    312.2,
           "Medium",     62.9,     77.3,    151.2,    163.3,    188.1,    191.8,    230.7,    139.5,      150,
            "Large",    259.6,    307.6,    398.7,    481.6,    566.2,    658.2,    741.8,    681.6,    772.3,
            "Total",    376.8,    460.6,      556,    651.9,    763.5,    859.2,    986.6,   1086.5,   1234.4 )

  
  
# Percentage from BI data								
percentage=tribble(
  ~`Size`, ~`2009/10`, ~`2010/11`, ~`2011/12`, ~`2012/13`, ~`2013/14`, ~`2014/15`, ~`2015/16`, ~`2016/17`, ~`2017/18`,
 "Small",  "14.4%",  "16.4%",  "20.8%",  "20.9%",  "22.2%",  "23.9%",  "26.4%",  "24.4%",    "25%",
           "Medium",  "16.7%",  "16.8%",  "16.0%",  "13.9%",  "12.1%",  "11.5%",  "10.9%",  "12.8%",    "12%",
            "Large",  "68.9%",  "66.8%",  "63.2%",  "65.2%",  "65.7%",  "64.6%",  "62.7%",  "62.7%",    "63%" )
```



```{r data table for actual Enterprise size, echo=F}
# Using data.table
datatable(actual_revenue_by.enterprise,caption = "Tax revenue by enterprise size:", extensions = 'Buttons',
options = list(dom='Bfrtip',                                                buttons=c('copy', 'csv', 'excel', 'pdf', 'print')))
```

```{r data table for % Enterprise size, echo=F}
datatable(percentage,caption = "Percentage tax revenue by enterprise size:",extensions = 'Buttons',options = list(dom='Bfrtip', buttons=c('copy', 'csv', 'excel', 'pdf', 'print')))
```


# Visualization of tax revenue by enterprise size

```{r tax_rev_by_enterprise_YB_Chart,echo=F}
viz1=actual_revenue_by.enterprise[1:3,] %>%  gather(key = 'Year',value = 'Revenue',-`Size`)

viz2=percentage[1:3,] %>% gather(key = 'Year',value = 'Percentage',-`Size`)
chart.table=bind_cols(viz1,viz2[,3]) %>% mutate(Year=factor(Year))

g <- ggplot(chart.table,aes(x=Year,y=Revenue, group=`Size`))+
 geom_point(size=1)+
  geom_line(aes(color=`Size`))+
  theme_bw()+
  labs(x='Fiscal year',y='Tax revenue (Rwf bn)', title='Tax revenue by enterprize size', fill='Enterprise size', caption = "Source: RRA data", tag= 'Fig.2')+
  theme(legend.position = 'top',axis.title.y = element_text(size = 12, face='bold'), axis.title.x =element_text(size = 12, face='bold'))

ggplotly(g)

```

# Tax revenue by enterprise type

```{r tax revenue by tax enterprise type FY2017_18,include=F}
# Grouping tax payers into enterprise type and changing the revenue into RWF billion

tax_rev1718_ET= tax_rev1718%>%select(1,2,5,28) %>% mutate_at(vars(4),~parse_number(.))%>% rename(Type=Ent.Type.Group) %>% mutate_at(vars(3),~as_factor(.))%>% 
  mutate(Type=fct_recode(Type, 'Private sector'='PRIVATE',
      'Public sector'='PUBLIC')) %>% 
  group_by(Type) %>% 
  summarise(`2017/18`=sum(Total,na.rm = T)) %>% mutate_at(vars(2),~round(./10^9,2))

```


```{r tax revenue by enterprise type FY2009_17, include=F}

# This data was from the analysis done before

tax_rev_2009_17_ET <- tribble(
  ~`Type`, ~`2009/10`, ~`2010/11`, ~`2011/12`, ~`2012/13`, ~`2013/14`, ~`2014/15`, ~`2015/16`, ~`2016/17`,
   "Private sector",   301.61,   382.06,   522.89,   511.94,   581.98,   677.98,   783.77,   854.37,
    "Public sector",    59.07,    81.73,   130.06,   152.43,   193.59,   204.65,   206.04,   230.47)

```
  
```{r tax revenue by tax enterprise type_YB,include=F}
# Combining tax revenue for FY2017/18 and FY2009/17
tax_rev_by_enterprise_type_YB <- bind_cols(tax_rev_2009_17_ET,tax_rev1718_ET[,2])     

# Calculating grand total
a=summarise_all(tax_rev_by_enterprise_type_YB[-1],sum);b=tibble(`Type`='Total')
Total=bind_cols(b,a)
tax_rev_by_enterprise_type_YB=bind_rows(tax_rev_by_enterprise_type_YB,Total)
```



```{r tax_rev_by_enterprisetype_final,include=F}

# We manipulated the data by using Percentage from the BI data `tax_rev_by_enterprise_type_YB` to the real collection data

# This is done in Excel sheet
							
actual_revenue_by.enterprise.type <- tribble(
  ~`Type`, ~`2009/10`, ~`2010/11`, ~`2011/12`, ~`2012/13`, ~`2013/14`, ~`2014/15`, ~`2015/16`, ~`2016/17`, ~`2017/18`,
  "Private sector", 315.1, 379.4, 445.3, 502.3, 572.9,   660, 781.2,  855.7,  970.4,
     "Public sector",  61.7,  81.2, 110.7, 149.6, 190.6, 199.2, 205.4,  230.8,    264,
             "Total", 376.8, 460.6,   556, 651.9, 763.5, 859.2, 986.6, 1086.5, 1234.4)

  
# Percentage from BI data								
percentage <- tribble(~`Type`, ~`2009/10`, ~`2010/11`, ~`2011/12`, ~`2012/13`, ~`2013/14`, ~`2014/15`, ~`2015/16`, ~`2016/17`, ~`2017/18`,
 "Private sector",  "83.6%",  "82.4%",  "80.1%",  "77.1%",  "75.0%",  "76.8%",  "79.2%",  "78.8%",  "78.6%", "Public sector",  "16.4%",  "17.6%",  "19.9%",  "22.9%",  "25.0%",  "23.2%",  "20.8%",  "21.2%",  "21.4%" )

```


```{r data table for actual Enterprise type, echo=F}
# Using data.table

datatable(actual_revenue_by.enterprise.type,caption = "Tax revenue by enterprise type:", extensions = 'Buttons',
options = list(dom='Bfrtip',                                                buttons=c('copy', 'csv', 'excel', 'pdf', 'print')))
```

```{r data table for % Enterprise type, echo=F}
datatable(percentage,caption = "Percentage tax revenue by enterprise type:",extensions = 'Buttons',options = list(dom='Bfrtip', buttons=c('copy', 'csv', 'excel', 'pdf', 'print')))
```


# Visualization of tax revenue by enterprise type

```{r tax_rev_by_enterprise.type_YB_Chart,echo=F}

viz1=actual_revenue_by.enterprise.type[1:2,] %>%  gather(key = 'Year',value = 'Revenue',-`Type`)

viz2=percentage[1:2,] %>% gather(key = 'Year',value = 'Percentage',-`Type`)

chart.table=bind_cols(viz1,viz2[,3]) %>%mutate(`Type`=as_factor(`Type`) %>% fct_relevel(.,'Public sector')) %>% mutate(Year=factor(Year))

g=ggplot(chart.table,aes(x=Year, y=Revenue,
  fill=`Type`))+
  geom_bar(stat = 'identity', position = position_dodge(), alpha=0.5, width = 0.9)+theme_bw()+labs(x='Fiscal year',y='Tax revenue (Rwf bn)', fill='Enterprise type', caption = "Source: RRA data", tag= 'Fig.3', title='Tax revenue by enterprise type')+
  theme(legend.position = 'top', axis.title.y = element_text(size = 12, face='bold'), axis.title.x =element_text(size = 12, face='bold'), axis.text.x = element_text(angle = 50, vjust = 0.3))+
scale_fill_manual(values = c("Private sector" = "green", "Public sector" = "red"), aesthetics = 'fill')+geom_text(aes(label=Percentage), vjust=1.2, size=2.5, colour='black',position = position_dodge(width = 0.9))
ggplotly(g)
```


```{r cleancache1, include=F}
knitr::clean_cache(clean = T)
```
                
# Part A: Tax revenue by Enterprise type description


```{r tax revenue by Individual vs Corporation Part A,cache=T, include=F}

# Appending Column `Year` into tax_rev_FY1718 dataset

tax_rev_FY1718=tax_rev1718 %>% mutate(Year='2017/18') %>% rename('T'=Total)


# Loading tax revenue data from FY2009 to FY2017

tax_rev_FY2009_17=read_csv('Dataset/tax_rev_2009_17.csv')


# Making the colnames of `tax_rev_FY1718` equal to `tax_rev_FY2009_17`

names(tax_rev_FY1718)=names(tax_rev_FY2009_17)

# Combining `tax_rev_FY2009_17` and `tax_rev_FY1718` together

tax_rev_FY2009_18=rbind(tax_rev_FY2009_17,tax_rev_FY1718)

# Selecting the columns that we wanted and grouping into enterprise type description
tax_rev_FY2009_18_IC_A=tax_rev_FY2009_18 %>% select(1,2,4,29,28)%>% mutate_at(vars(5),~parse_number(.)) %>% mutate_at(vars(3),~str_squish(.)) %>% 
  mutate_at(vars(3),~factor(.))%>% mutate(`Ent Type Desc`=fct_collapse(`Ent Type Desc`,
 `Association & Non Profit Org.`=c('ASSOCIATION', 'NON PROFIT ORG.','TRUST'),
 `Cooperative`   =  'COOPERATIVE',
  Embassy = c('CONSULAR OFFICE', 'EMBASSY', 'PRIVILEGED PERSONS'),     
    `Individual Enterprise`='INDIVIDUAL',
  `Joint Venture`='JOINT VENTURE',  
 `NGOs`=c('LOCAL NGOs','INTERNATIONAL NGOs'),
 `Others`='OTHERS',  
 `Partnership`='PARTNERSHIP',
 `Not Enterprise`=	'Not Enterprise',
 `Private Corporation`=c('Company','PRIVATE CORPORATION'),
 `Public Institution`=c('GOVERNMENT BODY','PUBLIC CORPORATION'))) 
 
tax_rev_FY2009_18_IC_A =tax_rev_FY2009_18_IC_A %>% group_by(`Ent Type Desc`,Year) %>% summarise(Total=sum(T, na.rm = T))

# Creating a pivot table for Year and Total
tax_rev_FY2009_18_IC_A=tax_rev_FY2009_18_IC_A %>% group_by(`Ent Type Desc`) %>% spread(Year,Total)
```

```{r tax revenue by Individual vs Corporation_YB Part A, include=F}

tax_rev_by_Ind_vs_Corp_YB_A <- tax_rev_FY2009_18_IC_A %>% mutate_at(vars(2:10),~(round(./10^9,2)))

tax_rev_by_Ind_vs_Corp_YB_A=tax_rev_by_Ind_vs_Corp_YB_A %>%ungroup(`Ent Type Desc`) %>% mutate(`Ent Type Desc`=parse_character(`Ent Type Desc`)) %>% rename(`Type desc.`=`Ent Type Desc`)


# Calculating Grand total
a=summarise_all(tax_rev_by_Ind_vs_Corp_YB_A[-1],sum,na.rm=T);b=tibble(`Type desc.`='Total')

Total=bind_cols(b,a)
tax_rev_by_Ind_vs_Corp_YB_A=bind_rows(tax_rev_by_Ind_vs_Corp_YB_A,Total)

```

```{r Individual vs Corporation_YB_final Part A, echo=F}
# We manipulated the data by using Percentage from the BI data `tax_rev_by_Ind_vs_Corp_YB_A` to the real collection data

# This is done in Excel sheet
							
actual_revenue_by.Ind_vs_Corp_A=tribble(
  ~`Type desc.`, ~`2009/10`, ~`2010/11`, ~`2011/12`, ~`2012/13`, ~`2013/14`, ~`2014/15`, ~`2015/16`, ~`2016/17`, ~`2017/18`,
  "Association & Non Profit Org.",      9.9,     15.6,     15.9,     15.2,     15.9,     16.6,     19.6,     20.1,     21.7,
            "Private Corporation",    235.3,    269.4,    309.3,    376.4,    434.7,    518.4,    634.7,    700.1,      817,
                        "Embassy",      0.2,      0.5,      0.4,      0.4,      0.4,      0.4,      0.4,      1.1,      1.1,
                    "Cooperative",      1.6,      3.1,      3.6,      5.3,      6.2,      6.6,      8.2,     10.5,     12.8,
             "Public Institution",     73.3,     99.8,    118.4,    144.6,    192.9,    202.1,      212,    236.6,      264,
          "Individual Enterprise",     27.1,     34.5,     65.3,     64.6,       62,     65.2,     66.6,     72.7,     67.4,
                           "NGOs",     14.2,     19.5,     16.8,       19,       21,     22.5,     25.8,     28.2,     32.5,
                  "Joint Venture",      0.6,      0.6,      0.3,      0.4,      0.2,      1.6,      0.5,      1.6,      1.2,
                 "Not Enterprise",      1.1,      1.8,     10.9,      9.7,      7.6,      7.5,     11.5,     12.7,     13.4,
                         "Others",     11.7,     14.2,     12.8,     14.8,     21.6,     17.5,      6.5,        2,      2.3,
                    "Partnership",      1.8,      1.6,      2.4,      1.5,        1,      0.8,      0.8,      0.8,      0.8,
                          "Total",    376.8,    460.6,      556,    651.9,    763.5,    859.2,    986.6,   1086.5,   1234.4 )

# Percentage from BI data								

percentage=tribble(~`Type desc.`, ~`2009/10`, ~`2010/11`, ~`2011/12`, ~`2012/13`, ~`2013/14`, ~`2014/15`, ~`2015/16`, ~`2016/17`, ~`2017/18`,
                           "Association & Non Profit Org.",   "2.6%",   "3.4%",   "2.9%",   "2.3%",   "2.1%",   "1.9%",   "2.0%",   "1.8%",   "1.8%",
                                     "Private Corporation",  "62.5%",  "58.5%",  "55.6%",  "57.7%",  "56.9%",  "60.3%",  "64.3%",  "64.4%",  "66.2%",
                                                 "Embassy",   "0.1%",   "0.1%",   "0.1%",   "0.1%",   "0.0%",   "0.0%",   "0.0%",   "0.1%",   "0.1%",
                                             "Cooperative",   "0.4%",   "0.7%",   "0.6%",   "0.8%",   "0.8%",   "0.8%",   "0.8%",   "1.0%",   "1.0%",
                                      "Public Institution",  "19.4%",  "21.7%",  "21.3%",  "22.2%",  "25.3%",  "23.5%",  "21.5%",  "21.8%",  "21.4%",
                                   "Individual Enterprise",   "7.2%",   "7.5%",  "11.7%",   "9.9%",   "8.1%",   "7.6%",   "6.8%",   "6.7%",   "5.5%",
                                                    "NGOs",   "3.8%",   "4.2%",   "3.0%",   "2.9%",   "2.7%",   "2.6%",   "2.6%",   "2.6%",   "2.6%",
                                           "Joint Venture",   "0.2%",   "0.1%",   "0.0%",   "0.1%",   "0.0%",   "0.2%",   "0.1%",   "0.1%",   "0.1%",
                                          "Not Enterprise",   "0.3%",   "0.4%",   "2.0%",   "1.5%",   "1.0%",   "0.9%",   "1.2%",   "1.2%",   "1.1%",
                                                  "Others",   "3.1%",   "3.1%",   "2.3%",   "2.3%",   "2.8%",   "2.0%",   "0.7%",   "0.2%",   "0.2%",
                                             "Partnership",   "0.5%",   "0.3%",   "0.4%",   "0.2%",   "0.1%",   "0.1%",   "0.1%",   "0.1%",   "0.1%")
                         
```



```{r data table for actual Enterprise type desc, echo=F}
# Using data.table
datatable(actual_revenue_by.Ind_vs_Corp_A,caption = "Tax revenue by enterprise type description:", extensions = 'Buttons',
options = list(dom='Bfrtip',                                                buttons=c('copy', 'csv', 'excel', 'pdf', 'print')))
```

```{r data table for % Enterprise type desc, echo=F}
datatable(percentage,caption = "Percentage tax revenue by enterprise type description:",extensions = 'Buttons',options = list(dom='Bfrtip', buttons=c('copy', 'csv', 'excel', 'pdf', 'print')))
```


# Part B: Tax revenue by Enterprise type description

```{r tax revenue by Individual vs Corporation Part B,cache=T, include=F}
# Selecting the column that we wanted and grouping into enterprise type description

tax_rev_FY2009_18_IC_B=tax_rev_FY2009_18 %>% select(1,2,4,29,28) %>% mutate_at(vars(5),~parse_number(.)) %>% 
  mutate_at(vars(4),~factor(.)) %>% 
  mutate(`Ent Type Desc`=recode_factor(`Ent Type Desc`,
           "INDIVIDUAL"='Individual enterprise',
            .default= 'Corporation')) 

tax_rev_FY2009_18_IC_B =tax_rev_FY2009_18_IC_B %>% group_by(`Ent Type Desc`,Year) %>% summarise(Total=sum(T, na.rm = T))

# Creating Pivot table for Year and Total

tax_rev_FY2009_18_IC_B=tax_rev_FY2009_18_IC_B %>% group_by(`Ent Type Desc`) %>% spread(Year,Total)
```

```{r tax revenue by Individual vs Corporation_YB_B, include=F}
# Convert the taxrevenue into RWF billion
tax_rev_by_Ind_vs_Corp_YB_B <- tax_rev_FY2009_18_IC_B %>% mutate_at(vars(2:10),~(round(./10^9,2)))

tax_rev_by_Ind_vs_Corp_YB_B=tax_rev_by_Ind_vs_Corp_YB_B %>%ungroup(`Ent Type Desc`) %>% mutate(`Ent Type Desc`=parse_character(`Ent Type Desc`)) %>% rename(`Type desc.`=`Ent Type Desc`)


# Calculating Grand total

a=summarise_all(tax_rev_by_Ind_vs_Corp_YB_B[-1],sum);b=tibble(`Type desc.`='Total')
Total=bind_cols(b,a)
tax_rev_by_Ind_vs_Corp_YB_B=bind_rows(tax_rev_by_Ind_vs_Corp_YB_B,Total)
```

```{r Individual vs Corporation_YB_final_B, echo=F}

# We manipulated the data by using Percentage from the BI data `tax_rev_by_Ind_vs_Corp_YB_A` to the real collection data

# This is done in Excel sheet

actual_revenue_by.Ind_vs_Corp_B=tribble(
  ~`Type desc.`, ~`2009/10`, ~`2010/11`, ~`2011/12`, ~`2012/13`, ~`2013/14`, ~`2014/15`, ~`2015/16`, ~`2016/17`, ~`2017/18`,
    "Individual enterprise",     27.1,     34.6,     65.3,     64.7,     62.2,     65.3,     66.7,     72.8,     67.4,
              "Corporation",    349.7,      426,    490.7,    587.2,    701.3,    793.9,    919.9,   1013.7,     1167,
                    "Total",    376.8,    460.6,      556,    651.9,    763.5,    859.2,    986.6,   1086.5,   1234.4)


# Percentage from BI data								

percentage=tribble(~`Type desc.`, ~`2009/10`, ~`2010/11`, ~`2011/12`, ~`2012/13`, ~`2013/14`, ~`2014/15`, ~`2015/16`, ~`2016/17`, ~`2017/18`,
  "Individual enterprise",   "7.2%",   "7.5%",  "11.8%",   "9.9%",   "8.1%",   "7.6%",   "6.8%",   "6.7%",   "5.5%",
            "Corporation",  "92.8%",  "92.5%",  "88.2%",  "90.1%",  "91.9%",  "92.4%",  "93.2%",  "93.3%",  "94.5%")
```

```{r data table for actual Enterprise type desc_B, echo=F}
# Using `DT` library ...

datatable(actual_revenue_by.Ind_vs_Corp_B,caption = "Tax revenue by enterprise type description:", extensions = 'Buttons',
options = list(dom='Bfrtip',                                                buttons=c('copy', 'csv', 'excel', 'pdf', 'print')))
```

```{r data table for % Enterprise type desc_B, echo=F}
# Using `DT` library ...

datatable(percentage,caption = "Percentage tax revenue by enterprise type description:",extensions = 'Buttons',options = list(dom='Bfrtip', buttons=c('copy', 'csv', 'excel', 'pdf', 'print')))
```

# Tax revenue by department

```{r tax revenue by department FY2009_18,include=F}

# The data were taking from Roy's Excel file and not BI datawarehouse

actual_revenue_by.dept=tribble(
  ~`Department`, ~`2009/10`, ~`2010/11`, ~`2011/12`, ~`2012/13`, ~`2013/14`, ~`2014/15`, ~`2015/16`, ~`2016/17`, ~`2017/18`,
              "DTD",    250.4,      300,      371,    449.3,    522.6,    594.3,    660.9,    729.3,    852.4,
          "Customs",      122,    157.3,    166.6,    180.3,    216.5,      239,    288.5,    312.6,    335.8,
        "Road Fund",     12.8,     16.6,     18.4,     22.3,     24.4,     25.9,     37.3,     44.6,     46.2,
            "Total",    385.2,    473.9,      556,    651.9,    763.4,    859.1,    986.7,   1086.5,   1234.4 )


# Percentage of the data								
percentage=tribble(~`Department`, ~`2009/10`, ~`2010/11`, ~`2011/12`, ~`2012/13`, ~`2013/14`, ~`2014/15`, ~`2015/16`, ~`2016/17`, ~`2017/18`,
                                         "DTD",  "65.0%",  "63.3%",  "66.7%",  "68.9%",  "68.5%",  "69.2%",  "67.0%",  "67.1%",  "69.1%",
                                     "Customs",  "31.7%",  "33.2%",  "30.0%",  "27.7%",  "28.4%",  "27.8%",  "29.2%",  "28.8%",  "27.2%",
                                   "Road Fund",   "3.3%",   "3.5%",   "3.3%",   "3.4%",   "3.2%",   "3.0%",   "3.8%",   "4.1%",   "3.7%")
                           
```

```{r data table for dept, echo=F}
# Using data.table
datatable(actual_revenue_by.dept,caption = "Tax revenue by department:", extensions = 'Buttons',
options = list(dom='Bfrtip',                                                buttons=c('copy', 'csv', 'excel', 'pdf', 'print')))
```

```{r data table for % department, echo=F}
# Using data.table
datatable(percentage,caption = "Percentage tax revenue by department:",extensions = 'Buttons',options = list(dom='Bfrtip', buttons=c('copy', 'csv', 'excel', 'pdf', 'print')))
```

# Visualization of tax revenue by department

```{r tax_rev_by_dept.chart,echo=F}

# unpivot columns as in Power Query in Excel

viz1=actual_revenue_by.dept[1:3,] %>%  gather(key = 'Year',value = 'Revenue',-`Department`)

viz2=percentage[1:3,] %>% gather(key = 'Year',value = 'Percentage',-`Department`)

chart.table=bind_cols(viz1,viz2[,3]) %>% mutate(`Department`=as_factor(`Department`) %>% fct_relevel(.,"Road Fund","Customs","DTD")) 

g=ggplot(chart.table,aes(x=Year, y=Revenue,
  fill=`Department`))+
  geom_bar(stat = 'identity', position = position_dodge(),  width = 0.9)+theme_bw()+labs(x='Fiscal year',y='Tax revenue (Rwf bn)', fill='Department', caption = "Source: RRA data", tag= 'Fig.4', title='Tax revenue by department')+
  theme(legend.position = 'bottom', axis.title.y = element_text(size = 12, face='bold'), axis.title.x = element_blank(), axis.text.x = element_text(angle = 50, vjust = 0.1))+
  geom_text(aes(label=Percentage), vjust=-1.3, size=2.4, colour='black',position = position_dodge(width = 0.9))
ggplotly(g)
```

```{r cleancache2, include=F}
knitr::clean_cache(clean = T)
```



This portal is powered by [Ezekiel Adebayo](https://bit.ly/gbganalyst).

